---
title: "Case Study"
author: "Joanna Duran"
date: "October 7, 2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Import files
```{r import, echo=TRUE}
beers<-read.csv("C:/Users/JEngi/Desktop/Doing data science/Beers.csv", stringsAsFactors = FALSE, header = TRUE)
breweries<-read.csv("C:/Users/JEngi/Desktop/Doing data science/Breweries.csv", stringsAsFactors = FALSE, header = TRUE)
```
```

```
*Beers present in each state
```
```
*Change column name to match so I can merge data

```{r name, echo=TRUE}
 names(breweries)[names(breweries)=="Brew_ID"]<-"Brewery_id"
```
*Merge Data
```{r merge, echo=TRUE}
beersnbrews<-merge(x=beers, y=breweries, by="Brewery_id", all=TRUE)
```

*Print first 6 and last 6 observations
```{r head, echo=TRUE}
head(beersnbrews,6)
```


```{r tail, echo=TRUE}
tail(beersnbrews,6)
```


```
```
ReportNumber of NAs in each column
```{r NAs, echo=TRUE}
colSums(is.na(beersnbrews))
```


Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r medalcohol, echo=TRUE}
medabv<-tapply(beersnbrews$ABV,beersnbrews$State, median)
barplot(medabv)
medibu<-tapply(beersnbrews$IBU,beersnbrews$State, median)
barplot(medibu)

```

Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r stateABV, echo=TRUE}
```

Summary statistics for the ABV variable.
```{r summaryABV, echo=TRUE}
summary(beersnbrews$ABV)
```

Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r bitvcont, echo=TRUE}
plot(beersnbrews$ABV,beersnbrews$IBU)
```
