---
title: "A Statistical Analysis of U.S. Beer Bitterness, Alcohol Content and State of Origin"
author: "Daniel Byrne, Joanna Duran, Matthew Norton"
date: "October 15, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.path = "_imgs/report-")
library(data.table)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(knitr)
library(pander)
library(reshape2)
library(usmap)
library(tidyr)
library(mosaic)
library(ggformula)
library(formattable)

# load data
beers = read.csv("beers.csv", header = TRUE,strip.white = TRUE)
colnames(beers) <-c("Beer_Name","Beer_ID","ABV","IBU","Brewery_ID","Style","Ounces")

beers$Brewery_ID <-as.character(beers$Brewery_ID)
beers$Beer_ID <-as.character(beers$Beer_ID)

breweries = read.csv("breweries.csv", header = TRUE,strip.white=TRUE)
colnames(breweries) <- c("Brewery_ID","Brewery_Name","City","State")
breweries$Brewery_ID <- as.character(breweries$Brewery_ID)
brewery_beer <- merge(breweries, beers, by=c("Brewery_ID"))
```
# Preface
Anheuser Busch inc, the Client, in coordination with Byrne, Durant Norton reseach LLP 
pursuant to Statement of Work number 29155, is evaluating statistical data on beers
for Anheuser Busch to use in Market Research.

Our team was presented the raw data on U.S. microbrewery beers provided by the market 
research department, and asked asked to analyze the data for patterns that could be used in developing new marketing opportunities.

# Introduction
This "Request For Comment", RFC, document  is being issued as an attempt to
stimulate a dialog on new market opportunituies in the ever growing microbrewery market.
In particular this RFC considers the raw beer statistics provided in the `beers.csv` and
`breweries.csv` provided to us by our Client and analyzes these data for possible trends
to lay the groundwork for possible aquisitions and marketing campaings.  

In particular we were asked to tally and vizualize the number of craft breweries by state,
summarize the statistical parameters of alcohol and bitterness of regional beers, and to 
identify any relationships of these parameters with each other and with the States in which
they were produced. 

# Background
According to Brewers Association for Small and Independent Craft Brewers the overall U.S. beer
volume sales were down 1% in 2017, whereas craft brewer sales continued to grow at a rate of 5%
by volume, reaching 12.7% of the U.S. beer market by volume. Specifically 'Craft' production grew
the most.

Specifically 'Craft' production grew the most. Retail dollar sales of craft brews increased 8%,
yar over year up to $26.0 billion. Craft brews now account for a quarter of the $111.4
billion U.S. beer market.

[https://www.brewersassociation.org/statistics/national-beer-sales-production-data/
](https://www.brewersassociation.org/statistics/national-beer-sales-production-data/)

# Analysis

### Microbreweries by State
```{r,echo=FALSE,message=FALSE}
brewery_cnt = data.frame(table(breweries$State))
colnames(brewery_cnt) <- c("abbr","total")
brewery <- merge(brewery_cnt,statepop, by = "abbr", all = FALSE)[,c("fips","full","abbr","total")]

## Subset to create regional brewery maps
northeast <-c("CT","DC","DE","MA","MD","ME","NH","NJ","NY","PA","RI","VT")
midwest <-c("ND","NE","SD","KS","MN","MO","IA","WI","MI","IL","IN","OH")
south <- c("AL","AR","FL","GA","KY","LA","MS","NC","OK","SC","TN","TX","VA","WV")
west <- c("AZ","CA","CO","ID","MT","NM","NV","OR","UT","WA","WY")
nc <- c("AK","HI")
brewery_northeast <- filter(brewery,abbr %in% northeast)
brewery_midwest <- filter(brewery,abbr %in% midwest)
brewery_south <- filter(brewery,abbr %in% south)
brewery_west <- filter(brewery,abbr %in% west)
brewery_nc <- filter(brewery,abbr %in% nc)

## Source mapfactory
if(!exists("new_map", mode = "function"))
    source("mapfactory.r")

## Create brewery maps
new_map(data = brewery, values = "total", include = c(),label = brewery$total, lines = "red") + 
  scale_fill_viridis_c(option = "magma", name = "Brewery Count") + 
  theme_void() + 
  ggtitle("US Breweries by State") + 
  theme(plot.title = element_text(hjust = 0.5))

new_map(data = brewery_northeast, values = "total", include = northeast, labels = TRUE, label = brewery_northeast$total, lines = "red") + 
  scale_fill_viridis_c() + 
  theme_void() + 
  ggtitle("Northeast Breweries by State") + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="none")

new_map(data = brewery_midwest, values = "total", include = midwest, labels = TRUE, label = brewery_midwest$total, lines = "red") +
  scale_fill_continuous(type = "viridis") + 
  theme_void() + 
  ggtitle("Midwest Breweries by State") + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="none")

new_map(data = brewery_south, values = "total", include = south, labels = TRUE, label = brewery_south$total, lines = "red") + 
  scale_fill_viridis_c(option = "plasma") + 
  theme_void() + ggtitle("Southern Breweries by State") + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="none")

new_map(data = brewery_west, values = "total", include = west, labels = TRUE, label = brewery_west$total, lines = "red") + 
  theme_void() + 
  ggtitle("Western Breweries by State") + 
  scale_fill_continuous() + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="none")

new_map(data = brewery_nc, values = "total", include = nc, labels = TRUE, label = brewery_nc$total, lines = "red") + 
  theme_void() + 
  ggtitle("Hawaii and Alaska Breweries by State") + 
  scale_fill_viridis_c(option = "viridis") + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="none")

```


### Merged Beer and Brewery Data Summary
```{r,echo=FALSE,message=FALSE}
brewery_beer <- merge(breweries, beers, by=c("Brewery_ID"))

# Merge head and tail subsets
top_bot_brewery <- rbind(
        select(head(brewery_beer),-c("Brewery_ID", "Beer_ID")),
        select(tail(brewery_beer),-c("Brewery_ID", "Beer_ID")))

# Format table
kable(top_bot_brewery, align = "l", col.names = c("Brewery","City","State","Beer","ABV","IBU","Style","Ounces"), row.names = FALSE) %>% 
  kable_styling(c("striped","condensed"), full_width = F, position = "left", font_size = 10) %>% 
  group_rows("Top Six Observations", 1,6, colnum = NULL, bold = T, label_row_css = "text-align: center") %>% 
  group_rows("Bottom Six Observations", 7,12, colnum = NULL, bold = T, label_row_css = "text-align: center")
```

### NA's Report
```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Count na, notna, null, and distinct values
t_brewery_na <- data.frame(add_rownames(data.frame(
  sapply(brewery_beer, function(x) sum(is.na(x))), 
  sapply(brewery_beer, function(x) sum(!is.na(x))),
  sapply(brewery_beer, function(x) sum(is.null(x))),
  sapply(brewery_beer, function(x) sum(n_distinct(x)))),
  "VALUE"))

colnames(t_brewery_na) <- c("Column","NA_totals","Not_NA_totals","Null_totals","Unique_totals")
kable(t_brewery_na, align = "l", caption = "NA, Not NA, Null and Unique Total Counts", row.names = FALSE) %>% 
  kable_styling(c("striped","condensed"), full_width = F, position = "left", font_size = 10)
```
### Statistical Analysis of Beer Variables by State
```{r, echo=FALSE,warning=FALSE}
ibuByState <- favstats(IBU~State,data=brewery_beer)
abvByState <- favstats(ABV~State,data=brewery_beer)

#order by median
ibuByState <- ibuByState[order(ibuByState$median,decreasing = TRUE),]
abvByState <- abvByState[order(abvByState$median,decreasing = TRUE),]

# chart and plot top 10 median IBU by State
gf_col(median~State,data=head(ibuByState,10), color=~median,fill =~median, title="Top 10 Median IBU by State")
kable(head(ibuByState,10), align = "l", row.names = FALSE, caption="Top 10 Median IBU by State") %>%
  kable_styling(c("striped","condensed"), full_width = F, position = "left", font_size = 10)

# chart and plot top 10 median ABV by State
gf_col(median~State,data=head(abvByState,10),color=~median,fill=~median, title="Top 10 Median ABV by State")
kable(head(abvByState,10), align = "l", row.names = FALSE, caption="Top 10 Median ABV by State") %>%
  kable_styling(c("striped","condensed"), full_width = F, position = "left", font_size = 10)

```

### State Maximums
```{r, echo=FALSE,warning=FALSE}
brewery_max <- merge(x=brewery_beer, y=statepop, by.x='State', by.y='abbr', all = FALSE)[,c("Brewery_Name","City","State","Beer_Name","ABV","IBU","full","fips")]

brewery_max$ABV_per <- percent(brewery_max$ABV)

abv_beer <- data.frame(brewery_max[which.max(brewery_max$ABV_per),])[,c("Brewery_Name","Beer_Name","ABV_per","City","State","full","fips")]

ibu_beer <- data.frame(brewery_max[which.max(brewery_max$IBU),])[,c("Brewery_Name","Beer_Name","IBU","City","State","full","fips")]

## Source mapfactory
if(!exists("abv_ibu_map", mode = "function"))
    source("mapfactory.r")

abv_ibu_map(data = abv_beer, values = "ABV_per", include = c(abv_beer$fips), label = (paste0(abv_beer$full,"\n",abv_beer$ABV_per)), lines = "red") + 
  scale_fill_continuous(na.value = "red", high = "blue") + 
  theme_void() + 
  ggtitle(paste0("Highest Alcohol(ABV) Beer","\n","Brewery: ",abv_beer$Brewery_Name,"\n","Beer: ",abv_beer$Beer_Name,"\n","Location: ",abv_beer$City,", ",abv_beer$State)) + 
  theme(plot.title = element_text(hjust = 0.5), legend.position="none")

abv_ibu_map(data = ibu_beer, values = "IBU", include = c(ibu_beer$fips), label = (paste0(ibu_beer$full,"\n",ibu_beer$IBU)), lines = "yellow") + 
  scale_fill_continuous(na.value = "yellow", high = "green") + 
  theme_void() + 
  ggtitle(paste0("Bitterest(IBU) Beer","\n","Brewery: ",ibu_beer$Brewery_Name,"\n","Beer: ",ibu_beer$Beer_Name,"\n","Location: ",ibu_beer$City,", ",ibu_beer$State)) +  
  theme(plot.title = element_text(hjust = 0.5), legend.position="none")
```

### Summary statistics for the ABV variable
```{r,echo=FALSE,warning=FALSE}
abvsummary <- favstats(~ABV,data=brewery_beer)
gf_boxplot(~ABV,data=brewery_beer,fill="yellow",outlier.color = "red")
kable(abvsummary, align = "l", row.names = FALSE, caption="Top 10 Median ABV by State") %>%
  kable_styling(c("striped","condensed"), full_width = F, position = "left", font_size = 10)
```

### Summary statistics for the IBU variable
```{r,echo=FALSE,warning=FALSE}
abvsummary <- favstats(~IBU,data=brewery_beer)
gf_boxplot(~IBU,data=brewery_beer,fill="green",outlier.fill="red")
kable(abvsummary, align = "l", row.names = FALSE, caption="Top 10 Median ABV by State") %>%
  kable_styling(c("striped","condensed"), full_width = F, position = "left", font_size = 10)
```

### Relationship between IBU and ABV
```{r,echo=FALSE,warning=FALSE}
# draw a point scatter plot showing the relationshp bewtween bitterness and alcohol content
gf_point(IBU~ABV,data=brewery_beer,color=~ABV-IBU)
```
